Opportunities for improvement in the care of adult trauma patients arriving in shock

### 2024-11-10 manuscript.Rmd
ändrade till passiv grej för theisis

The inclusion criteria for the trauma registry is a patient either admitted through trauma team activation or an already admitted patient with an Injury Severity Score (ISS) greater than nine. Every trauma patient (registered in the trauma registry) at Karolinska University Hospital is then included in a morbidity and mortality review process, which involves both individual case evaluations by specialized nurses and audit filters. Patients identified with a high potential for OFIs are discussed at multidisciplinary conferences. The identified OFIs are then categorized into broader areas. The presence or absence of OFIs is determined by consensus among all participants and is documented in the trauma care quality database.

>>>>

Inclusion in the trauma registry requires either admission through trauma team activation or presenting with an Injury Severity Score (ISS) greater than nine after admission to the Karolinska University Hopsital. Each trauma patient is then included in a morbidity and mortality review process, which involves both individual case evaluations by specialized nurses and audit filters. Patients identified with a high potential for OFIs are discussed at multidisciplinary conferences. The identified OFIs are then categorized into broad and detailed categories. The presence or absence of OFIs is determined by consensus among all participants and is documented in the trauma care quality database.

### 2024-11-10 thesis.Rmd
ändrade till passiv grej för theisis

We conducted a registry based retrospective cohort study, using data from the trauma registry and trauma care quality database at the Karolinska University Hospital in Solna.

@@

The trauma registry includes patients treated at Karolinska University Hospital in Solna, which treats all major trauma in the greater metropolitan area of Stockholm. We included patients registered between 2014 and 2023. The trauma care quality database is a subset of the trauma registry and includes patients selected for review.

@@

The inclusion criteria for the trauma registry is a patient either admitted through trauma team activation or an already admitted patient with an Injury Severity Score (ISS) greater than nine. Every trauma patient (registered in the trauma registry) at Karolinska University Hospital is then included in a morbidity and mortality review process, which involves both individual case evaluations by specialized nurses and audit filters. Patients identified with a high potential for OFIs are discussed at multidisciplinary conferences. The identified OFIs are then categorized into broader areas. The presence or absence of OFIs is determined by consensus among all participants and is documented in the trauma care quality database.

We included all patients in the trauma registry and Trauma care quality database. We excluded patients younger than 15 and/or were dead on arrival.

@@

The outcome was defined as the presence of at least one OFI, determined by the multidisciplinary M&M conference in the trauma care quality registry. An OFI can be various types of errors, which are categorized into: clinical judgment error, inadequate resources, delay in treatment, missed injury, inadequate protocols, preventable death and other errors.
The patients will be classified to different degrees of shock in parallel and separately, once with systolic blood pressure(SBP) and once with base excess (BE), regardless of their cause of shock. Based on these two parameters, the patients will be classified into four classes roughly based on the ATLS trauma shock classifications.
We included five other metrics to be adjusted for, due to potential for confounding. Pre-injury ASA and gender were categorical, meanwhile age, ISS and INR were kept continuous as they were directly from the register.

@@

All available data in trauma registry and trauma care quality database will be included.

@@

The BE parameter was divided into four classes with these cutoffs: Class I (above -2), Class II (-2 to -6), Class III (-6 to -10), and Class IV (below -10). Similarly, SBP was divided into four classes: Class I (above 110 mmHg), Class II (109-100 mmHg), Class III (99-90 mmHg), and Class IV (below 90 mmHg).
This classification follows the tenth edition ATLS hemorrhage classification, which does not specify exact values for any parameter other than BE. To solve this, we assigned numerical SBP values to each class based on findings from Eastridge et al. and Oyetunji et al., who redefined hypotension as 110 mmHg respective 90 mmHg dependent on age[@eastridge_hypotension_2007; @oyetunji_redefining_2011]. Therefore, we used 110 mmHg as the lower limit for (Class I: no shock) - (ATLS: normal SBP) and below 90 mmHg for (Class IV: clear shock) - (ATLS: clear hypotension), class 2 and 3 divided equally in between. This also aligns with the classification done by Mutschler et al.[@mutschler_critical_2013].

@@

Please note that ATLS classification defines class I for the BE parameter as (0 to -2) however, we choose to follow its original source[@mutschler_renaissance_2013], which defines class I as base deficit (Inverted BE) ≤ 2, with no lower limit (i.e. an upper limit for BE).

@@

The statistical analysis will be performed using R, a programming language and environment for statistical computing. We will present the types of OFI as percentage distributions and then visualize them in a bar chart. Unadjusted logistic regression will be used to determine the association between the OFI, and degree of shock defined separately by classifying BE and SBP roughly according to ATLS classification of hemorrhagic shock. Adjusted logistic regression will then incorporate other patient factors such as age, sex, preinjury ASA, INR, and ISS. It will be presented as odds ratios (OR) between the presence of OFI, and shock classes. The OR will be determined with 95% confidence intervals, and a significance level of 5% will be used. All statistical analysis will first be done on synthetic data and later implemented on the data collected from the trauma registry and the trauma care quality database to ensure objectivity. Missing data will be addressed by listwise deletion.

>>>>

A registry-based retrospective cohort study was conducted using data from the trauma registry and trauma care quality database at the Karolinska University Hospital in Solna.

@@

The trauma registry is compromised of patients treated at Karolinska University Hospital in Solna, which treats all major trauma in the greater metropolitan area of Stockholm. We included patients registered between 2014 and 2023. The trauma care quality database is a subset of the trauma registry and includes patients selected for review.

@@

Inclusion in the trauma registry requires either admission through trauma team activation or presenting with an Injury Severity Score (ISS) greater than nine after admission to the Karolinska University Hopsital. Each trauma patient is then included in a morbidity and mortality review process, which involves both individual case evaluations by specialized nurses and audit filters. Patients identified with a high potential for OFIs are discussed at multidisciplinary conferences. The identified OFIs are then categorized into broad and detailed categories. The presence or absence of OFIs is determined by consensus among all participants and is documented in the trauma care quality database.

All patients in the trauma registry and Trauma care quality database were included. Patients younger than 15 and/or were dead on arrival were excluded.

@@

The outcome was defined as the presence of at least one OFI, determined by the multidisciplinary M&M conference in the trauma care quality registry. An OFI can be various types of errors, which are categorized into: clinical judgment error, inadequate resources, delay in treatment, missed injury, inadequate protocols, preventable death and other errors.
Patients were classified into different shock severity categories in parallel and separately, once with systolic blood pressure (SBP) and once with base excess (BE), irrespective of shock cause. Based on these two parameters, the patients will be classified into four classes roughly based on the ATLS trauma shock classifications.
Five other metrics were included to be adjusted for, due to potential for confounding. Pre-injury ASA and gender were categorical, meanwhile age, ISS and INR were kept continuous as they were directly from the register.

@@

All available data in trauma registry and trauma care quality database were included.

@@

The BE parameter was divided into four classes with these cutoffs: Class I (above -2), Class II (-2 to -6), Class III (-6 to -10), and Class IV (below -10). Similarly, SBP was divided into four classes: Class I (above 110 mmHg), Class II (109-100 mmHg), Class III (99-90 mmHg), and Class IV (below 90 mmHg).
This classification follows the tenth edition ATLS hemorrhage classification, which does not specify exact values for any parameter other than BE. To solve this, numerical SBP values were assigned to each class based on findings from Eastridge et al. and Oyetunji et al., who redefined hypotension as 110 mmHg respective 90 mmHg dependent on age[@eastridge_hypotension_2007; @oyetunji_redefining_2011]. Therefore, 110 mmHg was set as the lower limit for (Class I: no shock) - (ATLS: normal SBP), and below 90 mmHg for (Class IV: clear shock) - (ATLS: clear hypotension), class 2 and 3 divided equally in between. This also aligns with the classification done by Mutschler et al.[@mutschler_critical_2013].

@@

Please note that ATLS classification defines class I for the BE parameter as (0 to -2) however, the original source[@mutschler_renaissance_2013] defines class I as base deficit (Inverted BE) ≤ 2, with no lower limit (i.e. an upper limit for BE). The original source classification was chosen for this study.

@@

The statistical analysis was performed using R, a programming language and environment for statistical computing. OFI subcategories was presented as percentage distributions and visualized in a bar chart. Unadjusted logistic regression was used to determine the association between the OFI, and degree of shock defined by BE and SBP classifications. Adjusted logistic regression incorporated other patient factors such as age, sex, preinjury ASA, INR, and ISS. It was presented as odds ratios (OR) between the presence of OFI, and shock classes. The OR was determined with 95% confidence intervals, and a significance level of 5% was used. All statistical analysis was first done on synthetic data and then later implemented on the data collected from the trauma registry and the trauma care quality database to ensure objectivity. Missing data was addressed by listwise deletion.
